---
title: "evo_ecomp_ecology"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read in the data

```{r}
library(ggplot2)
library(readr)
library(grid)
library(cowplot)
all_data <- read_csv("~/repos/ecology_of_evolutionary_computation/analysis/all_data.csv")
all_data$id <- as.integer(all_data$seed)
sharing_collatz <- subset(all_data, selection=="sharing")
sharing_collatz$sel_param <- as.factor(sharing_collatz$sel_param)
resource_collatz <- subset(all_data, selection=="resource")
resource_collatz$sel_param <- as.factor(resource_collatz$sel_param)
lexicase_collatz <- subset(all_data, selection=="lexicase")
lexicase_collatz$sel_param <- as.factor(lexicase_collatz$sel_param)

representative_collatz <- subset(all_data, (all_data$sel_param == 10 | all_data$selection != "sharing") & (all_data$sel_param == -1 | all_data$selection != "lexicase") & (all_data$sel_param == -1 | all_data$selection != "resource") )

square_data <- read_csv("~/repos/ecology_of_evolutionary_computation/analysis/square_data.csv")
square_data$id <- as.integer(square_data$seed)
sharing_square <- subset(square_data, selection=="sharing")
sharing_square$sel_param <- as.factor(sharing_square$sel_param)
resource_square <- subset(square_data, selection=="resource")
resource_square$sel_param <- as.factor(resource_square$sel_param)
lexicase_square <- subset(square_data, selection=="lexicase")
lexicase_square$sel_param <- as.factor(lexicase_square$sel_param)

representative_square <- subset(square_data, (square_data$sel_param == .1 | square_data$selection != "sharing") & (square_data$sel_param == 11 | square_data$selection != "lexicase") & (square_data$sel_param == 500 | square_data$selection != "resource") )

othello_data <- read_csv("~/repos/evo_comp_ecology/analysis/othello_data.csv")
sharing_othello <- subset(othello_data, selection=="sharing")
sharing_othello$sel_param <- as.factor(sharing_othello$sel_param)
resource_othello <- subset(othello_data, selection=="resource")
resource_othello$sel_param <- as.factor(resource_othello$sel_param)
lexicase_othello <- subset(othello_data, selection=="lexicase")
lexicase_othello$sel_param <- as.factor(lexicase_othello$sel_param)

representative_othello <- subset(othello_data, (othello_data$sel_param == 1 | othello_data$selection != "sharing"))

dow_data <- read_csv("~/repos/ecology_of_evolutionary_computation//analysis/dow_data.csv")
dow_data$selection <- as.factor(dow_data$selection)
dow_data$id <- as.integer(dow_data$seed)
sharing_dow <- subset(dow_data, selection=="sharing")
sharing_dow$sel_param <- as.factor(sharing_dow$sel_param)
resource_dow <- subset(dow_data, selection=="resource")
resource_dow$sel_param <- as.factor(resource_dow$sel_param)
lexicase_dow <- subset(dow_data, selection=="lexicase")
lexicase_dow$sel_param <- as.factor(lexicase_dow$sel_param)

representative_dow <- subset(dow_data, (dow_data$sel_param == 10 | dow_data$selection != "sharing"))

representative_collatz$problem <- "Collatz"
representative_dow$problem <- "Dow"
representative_square$problem <- "Square"

rep_all <- rbind(representative_collatz, representative_square, representative_dow)
```

```{r}
ggplot(data=subset(representative_collatz, representative_collatz$update == 990)) + geom_violin(aes(x=selection, y=phylogenetic_diversity)) + theme_classic()

```

Okay, that looks high for lexicase and resource and low for sharing and tournament. Makes sense.

```{r}

ggplot(data=subset(representative_collatz, representative_collatz$update == 990)) + geom_violin(aes(x=selection, y=mrca_depth)) + theme_classic()

```

Low for lexicase and resource is consistent with the previous graph.

Does fitness sharing threshold matter?

```{r}
ggplot(data=sharing_collatz) + stat_summary(aes(x=update, y=phylogenetic_diversity, group=sel_param, color=sel_param, fill=sel_param), fun.data ="mean_cl_boot", geom = "smooth") + theme_classic() + scale_x_continuous("Update") + scale_y_continuous()
```
Ah, yes it does. Looks like there's a clear optimum in the vicinity of 10. Graphing similar metrics (evolutionary distinctiveness, mean_pairwise distance, etc) is similar.

Does that pattern hold across non-phylogenetic diveristy?

```{r}
ggplot(data=sharing_collatz) + stat_summary(aes(x=update, y=phenotypic_entropy, group=sel_param, color=sel_param, fill=sel_param), fun.data ="mean_cl_boot", geom = "smooth") + theme_classic() + scale_x_continuous("Update") + scale_y_continuous()
```
Huh. Apparently not. That's important.

What about genotypic diversity?
```{r}
ggplot(data=sharing_collatz) + stat_summary(aes(x=update, y=diversity, group=sel_param, color=sel_param, fill=sel_param), fun.data ="mean_cl_boot", geom = "smooth") + theme_classic() + scale_x_continuous("Update") + scale_y_continuous()
```
Those are all pretty close together, but if any of them are higher than the others, it's not 10.

Let's do some quick stats.

```{r}
share_collatz_endpoints <- subset(sharing_collatz, sharing_collatz$update == 990)
kruskal.test(phenotypic_entropy ~ sel_param, data=share_collatz_endpoints)
pairwise.wilcox.test(share_collatz_endpoints$phenotypic_entropy, share_collatz_endpoints$sel_param, p.adjust.method = "bonferroni")
```
Okay, those are significantly different from each other.

```{r}
kruskal.test(phylogenetic_diversity ~ sel_param, data=share_collatz_endpoints)
pairwise.wilcox.test(share_collatz_endpoints$phylogenetic_diversity, share_collatz_endpoints$sel_param, p.adjust.method = "bonferroni")
```

10 is significantly different from everything else. So these orderings are actually different from each other.

And what about the differences for the Dow chemical problem?

```{r}
dow_endpoints <- subset(representative_dow, representative_dow$update==990)
kruskal.test(phylogenetic_diversity ~ selection, data=dow_endpoints)
pairwise.wilcox.test(dow_endpoints$phylogenetic_diversity, dow_endpoints$selection, p.adjust.method = "bonferroni")
```

Looks like lexicase is the only one that's significantly different from the others.

How about for lexicase?

```{r}
collatz_endpoints <- subset(representative_collatz, representative_collatz$update==990)
kruskal.test(phylogenetic_diversity ~ selection, data=collatz_endpoints)
pairwise.wilcox.test(collatz_endpoints$phylogenetic_diversity, collatz_endpoints$selection, p.adjust.method = "bonferroni")
```

Eco-EA Lexicase are different from the others but not each other.

Let's make a bunch of violin plots to check how various variables behave.

```{r}
ggplot(data=share_collatz_endpoints) + geom_violin(aes(x=sel_param, y=mean_pairwise_distance, fill=sel_param)) + theme_classic() +theme(legend.position = "none") + scale_x_discrete("Sharing threshold") + scale_y_continuous("Mean pairwise distance")
ggsave("../figs/sharing_pairwise_dist.png")

ggplot(data=share_collatz_endpoints) + geom_violin(aes(x=sel_param, y=phenotypic_entropy, fill=sel_param)) + theme_classic() +theme(legend.position = "none") + scale_x_discrete("Sharing threshold") + scale_y_continuous("Mean pairwise distance")
ggsave("../figs/sharing_phenotypic_entropy.png")

ggplot(data=subset(representative_collatz, representative_collatz$update == 990)) + geom_violin(aes(x=selection, y=phylogenetic_diversity, fill=selection)) + theme_classic() +theme(legend.position = "none") + scale_x_discrete("Selection scheme") + scale_y_continuous("Phylogenetic diversity")
ggsave("../figs/phylo_all.png")

ggplot(data=subset(representative_collatz, representative_collatz$update == 990)) + geom_violin(aes(x=selection, y=phenotypic_entropy, fill=selection)) + theme_classic() +theme(legend.position = "none") + scale_x_discrete("Selection scheme") + scale_y_continuous("Phylogenetic diversity")
ggsave("../figs/pheno_all.png")


p2 <- ggplot(data=representative_collatz) + stat_summary(aes(x=update, y=phenotypic_entropy, group=selection, color=selection, fill=selection), fun.data ="mean_cl_boot", geom = "smooth") + theme_classic() + scale_x_continuous("Generation") + scale_y_continuous("Phenotypic diversity") + theme(axis.title = element_text(size=18), axis.text = element_text(size=14), legend.text = element_text(size=14), legend.title = element_blank(), legend.position = "none") + scale_color_discrete(labels=c("Lexicase", "Eco-EA", "Sharing", "Tournament")) + scale_fill_discrete(guide="none")

p3 <- ggplot(data=representative_square) + stat_summary(aes(x=update, y=phylogenetic_diversity, group=selection, color=selection, fill=selection), fun.data ="mean_cl_boot", geom = "smooth") + theme_classic() + scale_x_continuous("Generation") + scale_y_continuous("Phylogenetic diversity") + theme(axis.title = element_text(size=18), axis.text = element_text(size=14), legend.text = element_text(size=14), legend.title = element_blank(), legend.position = "none") + scale_color_discrete(labels=c("Lexicase", "Eco-EA", "Sharing", "Tournament")) + scale_fill_discrete(guide="none")


p4 <- ggplot(data=representative_square) + stat_summary(aes(x=update, y=phenotypic_entropy, group=selection, color=selection, fill=selection), fun.data ="mean_cl_boot", geom = "smooth") + theme_classic() + scale_x_continuous("Generation") + scale_y_continuous("Phenotypic diversity") + theme(axis.title = element_text(size=18), axis.text = element_text(size=14), legend.text = element_text(size=14), legend.title = element_blank(), legend.position = "top") + scale_color_discrete(labels=c("Lexicase", "Eco-EA", "Sharing", "Tournament")) + scale_fill_discrete(guide="none")

```

Final code for phylogenetic and phenotypic diversity over time across problems:

```{r}
pleg <- ggplot(data=rep_all) + stat_summary(aes(x=update, y=phylogenetic_diversity, group=selection, color=selection, fill=selection), fun.data ="mean_cl_boot", geom = "smooth") + theme_classic() + scale_x_continuous("") + scale_y_continuous("Phylogenetic\ndiversity") + theme(axis.title = element_text(size=18), axis.text = element_text(size=12), legend.text = element_text(size=18), legend.title = element_blank(), legend.position = "bottom", strip.text = element_text(size=18)) + scale_color_discrete(labels=c("Lexicase", "Eco-EA", "Sharing", "Tournament")) + scale_fill_discrete(guide="none") + facet_wrap(~problem)

p2 <- ggplot(data=rep_all) + stat_summary(aes(x=update, y=phylogenetic_diversity, group=selection, color=selection, fill=selection), fun.data ="mean_cl_boot", geom = "smooth") + theme_classic() + scale_x_continuous("") + scale_y_continuous("Phylogenetic\ndiversity (1000s)", labels=function(x){x/1000}) + theme(axis.title = element_text(size=20), axis.text = element_text(size=12), legend.text = element_text(size=18), legend.title = element_blank(), legend.position = "none", strip.text = element_text(size=20)) + scale_color_discrete(labels=c("Lexicase", "Eco-EA", "Sharing", "Tournament")) + scale_fill_discrete(guide="none") + facet_wrap(~problem)

p3 <- ggplot(data=rep_all) + stat_summary(aes(x=update, y=phenotypic_entropy, group=selection, color=selection, fill=selection), fun.data ="mean_cl_boot", geom = "smooth") + theme_classic() + scale_x_continuous("Generation") + scale_y_continuous("Phenotypic\ndiversity") + theme(axis.title = element_text(size=20), axis.text = element_text(size=12), legend.text = element_text(size=18), legend.title = element_blank(), legend.position = "none", strip.text = element_text(size=20)) + scale_color_discrete(labels=c("Lexicase", "Eco-EA", "Sharing", "Tournament")) + scale_fill_discrete(guide="none") + facet_wrap(~problem)

pcol <- plot_grid(p2, p3, nrow=2, align="v")

l <- get_legend(pleg)
plot_grid(pcol, l, ncol = 1, rel_heights = c(1, .08))
ggsave("../figs/time_all.png", width = 10, height=6, scale=.8)

```

Final code for sharing thresholds over time:

```{r}
colnames(sharing_collatz)[32] = "Threshold"
legplot <- ggplot(data=sharing_collatz) + stat_summary(aes(x=update, y=phenotypic_entropy, group=Threshold, color=Threshold, fill=Threshold), fun.data ="mean_cl_boot", geom = "smooth") + theme_classic() + scale_x_continuous("Generation") + scale_y_continuous("Phenotypic diversity") + theme(axis.title = element_text(size=20), axis.text = element_text(size=12), legend.title = element_text(size=18), legend.text = element_text(size=18), legend.position = "bottom", strip.text = element_text(size=20)) + scale_color_discrete("Sharing theshold") + scale_fill_discrete(guide="none")
leg <- get_legend(legplot)
share_p1 <- ggplot(data=sharing_collatz) + stat_summary(aes(x=update, y=phenotypic_entropy, group=Threshold, color=Threshold, fill=Threshold), fun.data ="mean_cl_boot", geom = "smooth") + theme_classic() + scale_x_continuous("Generation") + scale_y_continuous("Phenotypic\ndiversity") + theme(axis.title = element_text(size=22), axis.text = element_text(size=14), legend.text = element_text(size=20), legend.title = element_blank(), legend.position = "none", strip.text = element_text(size=20))
share_p2 <- ggplot(data=sharing_collatz) + stat_summary(aes(x=update, y=phylogenetic_diversity, group=Threshold, color=Threshold, fill=Threshold), fun.data ="mean_cl_boot", geom = "smooth") + theme_classic() + scale_x_continuous("Generation") + scale_y_continuous("Phylogenetic\ndiversity (1000s)", labels=function(d){d/1000}) + theme(axis.title = element_text(size=22), axis.text = element_text(size=14), legend.text = element_text(size=20), legend.title = element_blank(), legend.position = "none", strip.text = element_text(size=20))

p_row <- plot_grid(share_p1, share_p2, align = "h")
plot_grid(p_row, leg, nrow = 2, rel_heights = c(1,.08))
ggsave("../figs/time_sharing.png", width = 18, height=6, scale=.7)

```

